AXIELL piskoviste

### 2023-05-17 _GEN_axi_Rhacks.Rmd
MUZBE: final modifications

@@

@@

@@

### Create a string of multiple column values with NAs

@@

# commands
dbname <- paste0(path_mdb, "bio_slovniky.accdb") # path to database
con <- RODBC::odbcConnectAccess2007(dbname)      # set the connection

@@

kmen <- RODBC::sqlFetch(con, "PRM_BioLib_080_Animalia_Kmen") # read specified table
```

>>>>

## I M P O R T / E X P O R T

### Read txt

read.table

```{r READ TABLE}
df <- read.table("D:/VURV/R_hacks/data/simple_table_4read.txt", header = TRUE, sep = ";")
```

### Read csv

read.csv : sep is always ","

```{r READ CSV}
df <- read.csv("D:/VURV/R_hacks/data/simple_table_4read.csv", header = TRUE, sep = ";")
```

### Read excel

```{r READ EXCEL}

library("readxl")
df <- read_excel("D:/VURV/R_hacks/data/simple_table_4read.xlsx" , sheet = 2, header = TRUE)
```

ISSUE: 'Warning: Expecting numeric in E1251 / R1251C5: got '41 (cm)'

SOLUTION: df <- read_excel(paste0(path_snd, "muzbe_rozmery_ptaci.xlsx"), sheet = 1, **col_types="text"**)

@@

### Remove last 35 columns

```{r DROP LAST X COLUMNS}

select(-mus_invc, -(last_col(offset = 34):last_col()))
```

@@

### str_trim pro cely df

```{r STR_TRIM DF}

df %>% 
  mutate(across(where(is.character), str_trim))
```

@@

### Unite

Creates a string of multiple column values with NAs.

@@

# general

dbname <- paste0("---cesta+jmeno-databaze.accdb") 
con <- RODBC::odbcConnectAccess2007(dbname)      
RODBC::sqlTables(con)
tab <- RODBC::sqlFetch(con, "---jmeno-tabulky---")

# old func example
dbname <- paste0(path_mdb, "bio_slovniky.accdb") # path to database
con <- RODBC::odbcConnectAccess2007(dbname)      # set the connection

@@

kmen <- RODBC::sqlFetch(con, "PRM_BioLib_080_Animalia_Kmen") # read specified table

```

### 2023-05-17 _GEN_axi_piskoviste.Rmd
MUZBE: final modifications

```{r}
require(tidyverse)
```

@@

# *GENERAL*

## X M L

### import

@@

## D E M Z

@@

>>>>

```{r}
require(tidyverse)

path_snd <- "C:/Users/krizova/Documents/R/data/_piskoviste/"
```

@@

# X M L

## import

@@

# D E M Z

@@

# ROZMERY PTACI

MUZBE  
Pro Danu wide -> long format + poznamky.

```{r ROZMERY PTACI}

library("tidyverse")
library("readxl")

# kolo 1

df <- read_excel(paste0(path_snd, "muzbe_rozmery_ptaci.xlsx"), sheet = 1, col_types="text")

install.packages("reshape2")
require(reshape2)

df_long <- melt(df, id.vars = c("MU_INVC", "měřil", "vážil"))

sort(unique(df_long$variable))

rozm_sna <- df_long %>% 
  group_by(MU_INVC) %>% 
  mutate(poradi = seq_along(value))

rozm_bezna <- rozm_sna %>% 
  drop_na(value) %>% 
  mutate(
    pozn = case_when(str_detect(variable, "Hmotnost|váha", negate = T)&!is.na(měřil) ~ paste("Měřil/a:", měřil),
                     str_detect(variable, "Hmotnost|váha")&!is.na(vážil) ~ paste("Vážil/a:", vážil),
                     TRUE ~ ""))

write.table(rozm_bezna, file = paste0(path_snd, "MUZBE_rozmery_ptaci_UPR.csv"),
            quote = T, row.names = F, 
            sep = ";", dec = ",", 
            na = "", fileEncoding="cp1250")

# desetinna cisla issue

dc <- rozm_bezna %>% 
  filter(str_detect(value, "\\...........*|,..........*"))

# kolo 2 - uspesne

df <- read.table(paste0(path_snd, "Pz_ZOO_Rozmery_Ptaci.txt"), header = T, sep = ";", check.names = F, fileEncoding="cp1250") %>%
  mutate(InvC = str_pad(InvC, 5, pad = "0"),
         MU_INVC = paste0("FM", InvC, "/Z")) %>% 
  relocate(MU_INVC, .after = InvC) %>% 
  select(-InvC, -Druh_CZ)

require(reshape2)

df_long <- melt(df, id.vars = c("MU_INVC", "id_meril", "id_vazil"))

sort(unique(df_long$variable))
df_long[df_long == ""] <- NA

rozm_bezna <- df_long %>% 
  drop_na(value) %>% 
  mutate(
    pozn = case_when(str_detect(variable, "vaha", negate = T)&!is.na(id_meril) ~ paste("Měřil/a:", id_meril),
                     str_detect(variable, "vaha")&!is.na(id_vazil) ~ paste("Vážil/a:", id_vazil),
                     TRUE ~ "")) %>% 
  filter(!value == "0")

write.table(rozm_bezna, file = paste0(path_snd, "MUZBE_rozmery_ptaci_UPR2.csv"),
            quote = T, row.names = F, 
            sep = ";", dec = ",", 
            na = "", fileEncoding="cp1250")
